Decorrelating Errors in Quantum Gates by Random Gate Synthesis

### 2018-08-02 decorrelation/decorrelation.tex
more references and changes

Steady progress has been made in the theory of quantum error correction, proving higher thresholds for increasingly general models of noise \cite{Aharonov2006}\needcite. These results show that quantum computation is feasible, however recent NISQ \cite{Preskill2018} devices have noise that is not only often above thresholds\needcite, but that also violates fundamental assumptions made by the models used in these results, such as Markovianity \cite{Kitaev1997} and independence of errors \needcite. With these assumptions violated, promises about system performance and correctness are difficult to make.

There are many ways in which device noise can be non-Markovian ways,\cite{Kelly2018, BlumeKohout2017}, even when the noise in a system is Markovian, many models make simplifying assumptions about the noise on that system. For instance Pauli channels are often used to model systems\needcite due to their classical simulability\needcite, despite their unrealistic form. Some authors have then taken the approach of using these easily-understood models to upper bound the error in a real system by \textit{honestly} representing them \cite{Puzzuoli2014}. This approach is correct and rigorous, but scales poorly.

Other authors have tried to address these problems \cite{Wallman2016, Campbell2017, Heim2016} by using circuit composition to remove correlated noise. Given a process that is generated by a sequence of channels with some unknown, potentially channel dependent error, coherent errors can be averaged away into incoherent error, at the cost of additional compilation steps, which may in general increase circuit depth and require more sophisticated classical preprocessing. Some of these routines have been demonstrated experimentally \cite{Ware2018}, and been shown to reduce the coherence of noise.

@@

In this example we intentionally over and under-calibrated four pulseshapes. For each of these pulseshapes, in addition to their BCS and the calibrated pulse, we then ran a randomized benchmarking experiment.\needcite The results are plotted in \ref{fig:rb}. In each case, our Clifford sequences were decomposed into RX($\frac{\pi}{2})$) and RY($\frac{\pi}{2})$) pulses.  In the superconducting qubit architecture used, these gates are implemented using the same pulse envelope definition. It was shown in \cite{Ball2016} that for a particular error model, non-Markovian, or \textit{quasi-static}, noise can result in more general Gamma Distributed noise, while Markovian noise, such as depolarizing noise, should result in Gaussian distributed fidelity estimates for each randomized benchmarking sequence depth. Our results are consistent with this, as demonstrated by the long tails in the four leftmost plots that vanish in the bottom right plot.

@@

\section{Conclusion}
We have shown numerically and experimentally that drawing from a collection of implementations of a gate with the correct probabilties can reduce coherent error by an order of magnitude, while increasing incoherent error by a smaller amount, at virtually no cost to gate fidelity. We have demonstrated that these approximate controls can be generated through optimal control, and that the minimization problem is tractable. In addition, we have shown that it is possible to perform the routine on existing quantum hardware. Future directions for this work include moving the random gate selection from a precompilation step onto runtime FPGA logic, investigating other optimzation routines such as CRAB \cite{Caneva2011} and GOAT\cite{Machnes2018}, and using more precise benchmarking routines such as GST\cite{BlumeKohout2017} to more quantitatively investigate the performance of these routines. The code used in this paper is available at \cite{decorrelating_errors}.

\section{Acknowledgements}
%It might be interesting to show a dhistogram of the pribabilities assigned to each control to show that they aren't all zero. and maybe

%$$ \epsilon, \delta \midtilde \sim \mathcal{N}(0, .001) $$
%$$\epsilon_0, \epsilon_1, \delta_0, \delta_1 \midtilde \sim \mathcal{N}(0, .001)$$
\bibliography{decorrelation.bib}
\end{document}

>>>>

Steady progress has been made in the theory of quantum error correction, proving higher thresholds for increasingly general models of noise \cite{Aharonov2006}. These results show that quantum computation is feasible, however recent NISQ \cite{Preskill2018} devices have noise that is not only often above thresholds, but that also violates fundamental assumptions made by the models used in these results, such as Markovianity \cite{Kitaev1997} and independence of errors\needcite. With these assumptions violated, promises about system performance and correctness are difficult to make.

There are many ways in which device noise can fail to be Markovian,\cite{Kelly2018, BlumeKohout2017}, and even when the noise \textit{is} Markovian, many models make simplifying assumptions about the structure of the noise. For instance, Pauli channels are often used to model systems\needcite due to their classical simulability\needcite, despite their unrealistic form. Some authors have then taken the approach of using these easily-understood models to upper bound the error in a real system by \textit{honestly} representing them \cite{Puzzuoli2014}. This approach is correct and rigorous, but scales poorly.

Other authors have tried to address these problems \cite{Wallman2016, Campbell2017, Heim2016} by using circuit compilation to remove correlated noise. Given a process that is generated by a sequence of channels with some unknown and potentially channel dependent error, coherent errors can be averaged away into incoherent error. Indeed, some of these routines have been demonstrated experimentally \cite{Ware2018}, and been shown to reduce the coherence of noise. However, this benefit comes at the cost of additional compilation steps, which may in general increase circuit depth and require more sophisticated classical preprocessing.

@@

In this example we intentionally over and under-calibrated four pulseshapes. For each of these pulseshapes, in addition to their BCS and the calibrated pulse, we then ran a randomized benchmarking experiment.\cite{Magesan2011} The results are plotted in \ref{fig:rb}. In each case, our Clifford sequences were decomposed into RX($\frac{\pi}{2})$) and RY($\frac{\pi}{2})$) pulses.  In the superconducting qubit architecture used, these gates are implemented using the same pulse envelope definition. It was shown in \cite{Ball2016} that for a particular error model, non-Markovian, or \textit{quasi-static}, noise can result in more general Gamma Distributed noise, while Markovian noise, such as depolarizing noise, should result in Gaussian distributed fidelity estimates for each randomized benchmarking sequence depth. Our results are consistent with this, as demonstrated by the long tails in the four leftmost plots that vanish in the bottom right plot.

@@

\section{Conclusion and Future Work}
We have shown numerically and experimentally that drawing from a collection of implementations of a gate with the correct probabilties can reduce coherent error by an order of magnitude, while increasing incoherent error by a smaller amount, at virtually no cost to gate fidelity. We have demonstrated that these approximate controls can be generated through optimal control, and that the minimization problem is tractable. In addition, we have shown that it is possible to perform the routine on existing hardware. The code used to generate the numerical results in this paper is available at \cite{decorrelating_errors}.

Future directions for this work include moving the random gate selection from a precompilation step to a runtime logic, investigating other optimzation routines such as CRAB \cite{Caneva2011} and GOAT\cite{Machnes2018}, and using more precise benchmarking routines such as GST\cite{BlumeKohout2017} to more quantitatively investigate the performance of these routines.

In addition, the numerical work in the paper assumed access to a model of the system. In general, an experimenter may not have a model readily available to describe the system, maybe due to on-chip crosstalk, or even if they do have access to such a model, it might be computationally intractable to simulate. In both of these situations, numerical generation of candidate pulseshapes and the minimization become infeasible. One approach would to use \textit{in situ} optimal control techniques \cite{Wu2018, Kelly2014, Ferrie2015} to generate candidate controls, and then use an optimizer like Nealder-Mead to minimize the magnitude of off-diagonal terms as measured by simple pulse sequences that amplify those elements of the process matrix.\needcite

\section{Acknowledgements}

\bibliography{decorrelation.bib}
\end{document}
